### 2020-12-17 paper/localpsf.tex
editing weighting functions section to be infinite dimensional

It is important to note that functions in $H^2$ only have pointwise values if the spatial dimension is $d=1$, $2$, or $3$. Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer. When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used.

\begin{defn}
	\label{def:stiffness_mass}
	We define $\mathcal{K}:H^2(\Omega)\rightarrow H^{-2}(\Omega)$ and $\mathcal{M}:L^2(\Omega) \rightarrow L^2(\Omega)'$ to be the following stiffness and mass linear operators:
	\begin{align*}
	\left(\mathcal{K}u\right)(v) :=& \int_\Omega \nabla u \cdot \nabla v ~dx \\
	\left(\mathcal{M}u\right)(v) :=& \int_\Omega u v ~dx,
	\end{align*}
	and $\mathcal{P}:H^2(\Omega)\rightarrow \mathbb{R}^r$ to be the following pointwise observation operator:
	\begin{equation*}
	(\mathcal{P}u)_i = u(x_i)
	\end{equation*}
	for $i=1,\dots,r$.
\end{defn}

\begin{prop}
	\label{prop:equivalent_writing_of_weighting_optimization}
	Let $f \in \mathbb{R}^r$ be arbitrary. The optimization problem
	\begin{equation}
	\label{eq:generic_optimization_initial}
	\begin{aligned}
	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r
	\end{aligned}
	\end{equation}
	may be rewritten as follows:
	\begin{equation}
		\label{eq:generic_optimization}
		\begin{aligned}
		\min_{u \in H^2(\Omega)} &\quad \frac{1}{2} \left(\mathcal{K}\mathcal{M}^{-1}\mathcal{K}u\right)(u) \\
		\text{such that} &\quad 
		\mathcal{P} u = f.
		\end{aligned}
	\end{equation}
\end{prop}

\begin{prop}
	There exists a unique solution to optimization problem \eqref{eq:generic_optimization}. The solution to \eqref{eq:generic_optimization}, $u^*$, and the adjoint variable for enforcing the constraint, $\lambda^*$, solve the following Karun-Kush-Tucker (KKT) system:
	\begin{equation}
	\label{eq:weighting_kkt}
		\begin{bmatrix}
			\mathcal{K}\mathcal{M}^{-1}\mathcal{K} & \mathcal{P}^T \\
			\mathcal{P} & 0
		\end{bmatrix}
		\begin{bmatrix}
			u^* \\ \lambda^*
		\end{bmatrix}
		=
		\begin{bmatrix}
			0 \\ f
		\end{bmatrix},
	\end{equation}
	where we use block-matrix notation to denote block-operators.
\end{prop}

@@

\begin{proof}
	By linearity of $\mathcal{T}$ (Corollary \ref{cor:linearity_of_solution_operator}), we have
	\begin{equation*}
		\sum_{i=1}^r w_i(x) = \sum_{i=1}^r \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^r e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
	\end{equation*}
	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one. The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function (i.e., $C(x_i) = 1$ for $i=1,\dots,r$,  and $\frac{1}{2}\|\Delta C\|^2 = 0$).
\end{proof}

@@

\begin{proof}
	Let
	\begin{equation*}
		\mathbf{v} := K \mathbf{w}.
	\end{equation*}
	Since the null-space of the Laplacian with pure Neumann boundary conditions is the set of all constant functions, we have
	\begin{equation}
	\label{eq:definition_of_w}
	\mathbf{w} = K^+ \mathbf{v} + \alpha \mathbf{c}
	\end{equation}
	for some scalar $\alpha$.

	Writing optimization problem \eqref{eq:discrete_poisson_optimization_1} in terms of $\mathbf{v}$ and $\alpha$ instead of $\mathbf{w}$ yields the following equivalent optimization problem:
	\begin{equation}
	\label{eq:discrete_poisson_optimization_2}
	\begin{aligned}
	\min_{\mathbf{v}, \alpha} &\quad \frac{1}{2}\mathbf{v}^T M^{-1} \mathbf{v} \\
	\text{such that} &\quad EK^+ \mathbf{v} + \alpha \mathbf{1} = \mathbf{d}.
	\end{aligned}
	\end{equation}
	The Lagrangian for optimization problem \eqref{eq:discrete_poisson_optimization_2} is
	\begin{equation*}
		\mathcal{L} = \frac{1}{2}\mathbf{v}^T M^{-1} \mathbf{v} + \boldsymbol{\lambda}^T\left(EK^+ \mathbf{v} + \alpha \mathbf{1} - \mathbf{d}\right).
	\end{equation*}
	The solution to this optimization problem is the stationary point of the Lagrangian, i.e., the point at which the gradient of the Lagrangian is zero:
	\begin{equation*}
		0 = \nabla \mathcal{L} = 
		\begin{bmatrix}
			\nabla_\mathbf{v} \mathcal{L} \\
			\nabla_{\boldsymbol{\lambda}} \mathcal{L} \\
			\nabla_\alpha \mathcal{L}
		\end{bmatrix} 
		= \begin{bmatrix}
			M^{-1} \mathbf{v} + K^+ E^T \boldsymbol{\lambda} \\
			EK^+\mathbf{v} + \alpha \mathbf{1} - \mathbf{d} \\
			\mathbf{1}^T \boldsymbol{\lambda}
		\end{bmatrix}.
	\end{equation*}
	This equation may be rewritten as the following block $3 \times 3$ linear system:
	\begin{equation*}
		\begin{bmatrix}
			M^{-1} & K^+ E^T & 0 \\
			EK^+ & 0 & \mathbf{1} \\
			0 & \mathbf{1}^T & 0
		\end{bmatrix}
		\begin{bmatrix}
			\mathbf{v} \\ \boldsymbol{\lambda} \\ \alpha
		\end{bmatrix}
		=
		\begin{bmatrix}
			0 \\ \mathbf{d} \\ 0
		\end{bmatrix}.
	\end{equation*}
	Performing block Gaussian elimination on this system allows us to reduce the system to the following block triangular form
	\begin{equation*}
		\begin{bmatrix}
			M^{-1} & K^+ E^T & 0 \\
			& -S & \mathbf{1} \\
			0 & 0 & \mathbf{1}^T S^{-1} \mathbf{1}
		\end{bmatrix}
		\begin{bmatrix}
			\mathbf{v} \\ \boldsymbol{\lambda} \\ \alpha
		\end{bmatrix}
		=
		\begin{bmatrix}
			0 \\ \mathbf{y} \\ \mathbf{1}^T S^{-1} y
		\end{bmatrix}
	\end{equation*}
	where we recall that $S := \Psi^T M \Psi = EK^+ M K^+ E^T$. From this block triangular system, we read off the solution as:
	\begin{align*}
		\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
		\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right) \\
		\mathbf{v} =& -M K^+ E^T \boldsymbol{\lambda}
	\end{align*}
	Substituting these results into the definition of $\mathbf{w}$ in \eqref{eq:definition_of_w}, we have
	\begin{equation*}
		\mathbf{w} = -K^+ M K^+ E^T \boldsymbol{\lambda} + \alpha \mathbf{c} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
	\end{equation*}
	as required.
\end{proof}

>>>>

It is important to note that functions in $H^2$ only have well-defined pointwise values if the spatial dimension is $d=1$, $2$, or $3$. Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer. When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used.

\begin{defn}
	\label{def:stiffness_mass}
	We define $\mathcal{K}:H^2(\Omega)\rightarrow L^2(\Omega)'$ and $\mathcal{M}:L^2(\Omega) \rightarrow L^2(\Omega)'$ to be the following linear operators:
	\begin{align*}
	\left(\mathcal{K}u\right)(v) :=& \int_\Omega v \Delta u ~dx \\
	\left(\mathcal{M}u\right)(v) :=& \int_\Omega u v ~dx,
	\end{align*}
	and $\mathcal{P}:H^2(\Omega)\rightarrow \mathbb{R}^r$ to be the following pointwise observation operator:
	\begin{equation*}
	(\mathcal{P}u)_i = u(x_i)
	\end{equation*}
	for $i=1,\dots,r$.
\end{defn}

\begin{prop}
	\label{prop:equivalent_writing_of_weighting_optimization}
	Let $f \in \mathbb{R}^r$ be arbitrary. The optimization problem
	\begin{equation}
	\label{eq:generic_optimization_initial}
	\begin{aligned}
	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r
	\end{aligned}
	\end{equation}
	may be rewritten as follows:
	\begin{equation}
		\label{eq:generic_optimization}
		\begin{aligned}
		\min_{u \in H^2(\Omega)} &\quad \frac{1}{2} \left(\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K}u\right)(u) \\
		\text{such that} &\quad 
		\mathcal{P} u = f.
		\end{aligned}
	\end{equation}
\end{prop}

\begin{prop}
	There exists a unique solution to optimization problem \eqref{eq:generic_optimization}. The solution to \eqref{eq:generic_optimization}, $u^*$, and the adjoint variable for enforcing the constraint, $\lambda^*$, solve the following Karun-Kush-Tucker (KKT) system:
	\begin{equation}
	\label{eq:weighting_kkt}
		\begin{bmatrix}
			\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K} & \mathcal{P}^T \\
			\mathcal{P} & 0
		\end{bmatrix}
		\begin{bmatrix}
			u^* \\ \lambda^*
		\end{bmatrix}
		=
		\begin{bmatrix}
			0 \\ f
		\end{bmatrix},
	\end{equation}
	where we use block-matrix notation to denote block-operators.
\end{prop}

@@

\begin{proof}
	By linearity of $\mathcal{T}$ (Corollary \ref{cor:linearity_of_solution_operator}), we have
	\begin{equation*}
		\sum_{i=1}^r w_i(x) = \sum_{i=1}^r \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^r e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
	\end{equation*}
	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one. The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function.
\end{proof}

@@

\begin{proof}
	Let
	\begin{equation*}
		v := \mathcal{K} w_i,
	\end{equation*}
	and for the remainder of this proof let $e := e_i$.
	Since the null-space of the Laplacian with pure Neumann boundary conditions is the set of all constant functions, we have
	\begin{equation}
	\label{eq:definition_of_w}
	w_i = \mathcal{K}^+ v + \alpha C
	\end{equation}
	for some scalar $\alpha$.

	Writing optimization problem \eqref{eq:wi_optimization_problem} in terms of $v$ and $\alpha$ instead of $w_i$ yields the following equivalent optimization problem:
	\begin{equation}
	\label{eq:discrete_poisson_optimization_2}
	\begin{aligned}
	\min_{v \in L^2(\Omega)', \alpha \in \mathbb{R}} &\quad \frac{1}{2} v\left(\mathcal{M}^{-1} v\right) \\
	\text{such that} &\quad \mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} = e.
	\end{aligned}
	\end{equation}
	The Lagrangian for optimization problem \eqref{eq:discrete_poisson_optimization_2} is
	\begin{equation*}
		\mathcal{L} = \frac{1}{2}v\left( \mathcal{M}^{-1} v\right) + \lambda^T\left(\mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} - e\right).
	\end{equation*}
	The solution to this optimization problem is the stationary point of the Lagrangian, i.e., the point at which the gradient of the Lagrangian is zero:
	\begin{equation*}
		0 = \nabla \mathcal{L} = 
		\begin{bmatrix}
			\nabla_{v} \mathcal{L} \\
			\nabla_{\lambda} \mathcal{L} \\
			\nabla_\alpha \mathcal{L}
		\end{bmatrix} 
		= \begin{bmatrix}
			\mathcal{M}^{-1} v + \mathcal{K}^+ \mathcal{P}^T \lambda \\
			\mathcal{P}\mathcal{K}^+v + \alpha \mathbf{1} - e \\
			\mathbf{1}^T \lambda
		\end{bmatrix}.
	\end{equation*}
	This equation may be rewritten as the following block $3 \times 3$ linear system:
	\begin{equation*}
		\begin{bmatrix}
			\mathcal{M}^{-1} & \mathcal{K}^+ \mathcal{P}^T & 0 \\
			\mathcal{P}\mathcal{K}^+ & 0 & \mathbf{1} \\
			0 & \mathbf{1}^T & 0
		\end{bmatrix}
		\begin{bmatrix}
			v \\ \lambda \\ \alpha
		\end{bmatrix}
		=
		\begin{bmatrix}
			0 \\ e \\ 0
		\end{bmatrix}.
	\end{equation*}
	Performing block Gaussian elimination on this system allows us to reduce the system to the following block triangular form
	\begin{equation*}
		\begin{bmatrix}
			\mathcal{M}^{-1} & \mathcal{K}^+ \mathcal{P}^T & 0 \\
			& -S & \mathbf{1} \\
			0 & 0 & \mathbf{1}^T S^{-1} \mathbf{1}
		\end{bmatrix}
		\begin{bmatrix}
			v \\ \lambda \\ \alpha
		\end{bmatrix}
		=
		\begin{bmatrix}
			0 \\ e \\ \mathbf{1}^T S^{-1} e
		\end{bmatrix}
	\end{equation*}
	where we recall that $S := \Psi^T \mathcal{M} \Psi = \mathcal{P}\mathcal{K}^+ \mathcal{M} \mathcal{K}^+ \mathcal{P}^T$. From this block triangular system, we read off the solution as:
	\begin{align*}
		\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
		\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right) \\
		\mathbf{v} =& -M K^+ E^T \boldsymbol{\lambda}
	\end{align*}
	Substituting these results into the definition of $\mathbf{w}$ in \eqref{eq:definition_of_w}, we have
	\begin{equation*}
		\mathbf{w} = -K^+ M K^+ E^T \boldsymbol{\lambda} + \alpha \mathbf{c} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
	\end{equation*}
	as required.
\end{proof}
